Synthesizing Symmetric Lenses

### 2019-02-12 papers/SymmetricLenses/main.tex
add main and macros fixes appendix and other stuff

{\em Lenses} are programs that can be run both ``front to back'' and ``back
to front,'' allowing updates to both input and output data to be transferred
in both directions.
Since their introduction by~\citet{Focal2005-long}, lenses have been extensively
studied, extended, and applied.  Recent work~\cite{optician} has also demonstrated how
techniques from {\em type-directed program synthesis} can be used to
efficiently synthesize a simple class of lenses---so-called {\em
  bijective lenses} over string data---given a pair of types (regular
expressions) and a small number of examples.  \bcp{Why is the next line not
  indented?} \afm{I have truly no idea, deferring until I get a stackoverflow answer.}

We show how to extend this synthesis algorithm to a much wider class of
lenses, which we call {\em simple symmetric lenses}, including not only all
of the bijective lenses and the more widely used ``asymmetric''
lenses~\cite{Focal2005-long}, but also a rich subset of the full-blown
``symmetric lenses'' proposed by \citet{symmetric-lenses}.

@@

In summary, our contributions are:
\begin{itemize}
%
\item We define {\em simple symmetric lenses}, a subclass of symmetric
lenses that do not maintain any hidden state (\S\ref{sec:overview}).  
%
We present a collection of combinators for building simple symmetric lenses and
describe how to calculate their likelihoods (\S\ref{sec:ssl}).
\item We give an algorithm for synthesizing lenses built from these combinators,
  using a novel application of stochastic regular expressions to guide the
  search process (\S\ref{sec:synthesis}) and prove that our metrics have deep
  semantic meaning.
%
\item We extend the Boomerang implementation with simple symmetric lenses,
  demonstrating that they integrate smoothly with other lens extensions such as
  quotient lenses~\cite{quotientlenses} and matching
  lenses~\cite{matchinglenses}. We evaluate our implementation, the effects of
  various optimizations, and our inclusion of relevance annotations on a set of
  48 lens synthesis benchmarks drawn from data cleaning, view maintenance, and
  file synchronization tasks. The system can synthesize simple symmetric lenses
  for all of the benchmarks in under 30 seconds (\S\ref{sec:evaluation}).
%
\item We show that simple symmetric lenses fall between asymmetric
lenses~\cite{Focal2005-long} and full symmetric
lenses~\cite{symmetric-lenses} in expressiveness.  Indeed, the
class of simple lenses can be characterized semantically as the subset of
full symmetric lenses for which a simple ``forgetfulness'' property holds
(\S\ref{sec:related}).
\end{itemize}
Along the
way, we show that {\em star-semiring equivalences} on regular expressions
remain valid when extended to stochastic regular expressions, preserving
probability distributions as well as languages (\S\ref{sec:sre}).
We close with related work (\S\ref{sec:related}) and
concluding thoughts (\S\ref{sec:conc}).
Elided proofs can be found in the appendices\bcp{The appendices are pretty
  hideous and hard to navigate!!}
\ifappendices\else included with the supplemental material\fi.

@@

  \setbox0=\hbox{%
    \begin{minipage}{3.5in}
\begin{lstlisting}[
stringstyle=\upshape\sffamily
]
Name,Company
Jane Doe,Healthcare Inc.
John Public,Insurance Co.
\end{lstlisting}
    \end{minipage}
  }
  \savestack{\listingB}{\box0}

@@

  \setbox0=\hbox{%
    \begin{minipage}{3.5in}
\begin{lstlisting}
let company = (name . ("Co." | "Inc." | "Ltd.")) | "UNK" in
let emp_ins = name . " " . name "," company in
let header = "Name,Company" in
let emp_insurance = header . ("\n" . emp_ins)$^*$
\end{lstlisting}
    \end{minipage}
  }
  \savestack{\listingD}{\box0}

  \centering
  \resizebox{\columnwidth}{!}{%
    \begin{tabular}{|>{\columncolor{vlightyellow}}c|>{\columncolor{vlightgreen}}c|}
      \hline
      Management's data & HR's data \\
      \hline
      \listingA & \listingB \\
      \hline
      \cellcolor{lightyellow}Management's type & 
      \cellcolor{lightgreen} HR's type\\
      \hline
      \cellcolor{lightyellow}\listingC & \cellcolor{lightgreen}\listingD \\
      \hline
    \end{tabular}
  }
  \caption{Hypothetical example data files and corresponding regular expressions
  used by management and HR at a US company to represent employee salaries and health
  insurance providers, respectively. \bcp{The colors are a bit loud,
    especially since there is so little color elsewhere.}
  %
  \bcp{Suggest replacing ``Name,Company'' with ``First Last, Company'' (too
    many ``name''s flying around)}
  %
  \bcp{More substantially, thinking about this example makes me wonder about
    a whole different way of specifying simple symmetric lenses: rather than
    just giving corresponding source and target pairs, could we give triples
    of the form ``new source / old target / new target'' and ``new target /
    old source / new source''?  I know these would be a bit more work to
    write, but they would also give the synthesis procedure way more
    information to work with---it seems to me it should be able to do much
    less guessing about where things are disconnected.  Would this allow us
    to drop all the information theory?  Or would we still need it?  (In
    retrospect, this seems like a fairly obvious thing to try, so it seems
    important to try it and report on how it performs relative to what we
    propose here...)}\afm{Oh sorry, yes we can write such examples.  I personally found that typically the same information could be provided through a series of 2 creates (aka: if we want a disconnect: creater a = b; creater a' = b:  if we want an identity: creater a = a; creater a' = a').  However, we also permit putr and putl examples of the proposed form in the tool 
Much like other aspects of the work, any situation with examples ultimately only constrains the subcomponents of the lens corresponding to that specific parse tree (aka: if I give an example about A in A|B, it only constrains "A").  I don't think it removes the need for the information theory, as all the other non-constrained subcomponents of the lens can still be "anything."
}  }
  \label{fig:minimized-representations}
\end{figure}

@@

\begin{theorem}
  If $\Regex \SSREquiv \RegexAlt$ then $\ProbabilityOf{\Regex}{\String} =
  \ProbabilityOf{\RegexAlt}{\String}$, for all strings $\String \in \LanguageOf{\Regex}$.
\end{theorem}
\begin{proof}
  See the appendix%
\ifappendices\else{} included with the supplemental
material\fi.  \bcp{It took me quite a while to find it.  We
  should give an actual pointer to a specific theorem in the appendix.}
\end{proof}

@@

We now give typing rules and semantics for (some of) these combinators. We do not
include the semantics for combinators equivalent to those presented in prior
work~\cite{Boomerang}. Furthermore, we present
just \CreateR and \PutR in cases where \CreateL and \PutL are symmetric;
full details can be found in an appendix\bcp{where,
  exactly??}\ifappendices\else{} (in the supplemental materials)\fi. For
simplicity of presentation, we use unambiguous regular expressions
everywhere, even though the identity and disconnect lenses can actually be defined over ambiguous
regular expressions; we believe relaxing
this restriction is unproblematic.

@@

The details of these definitions are syntactically tedious, but not
intellectually difficult. We elide them here but include them in the appendix.\bcp{Where?}

@@

\subsection{Optimizations and Relevance Annotations}
\label{subsec:optimizations}

@@

\paragraph*{Relevance Annotations}\bcp{Wonder whether this should be its own
top-level section...?}\afm{I'm not sure, any other thoughts? To me it is
inherently an optimization, though it's a more important and interesting one
than others}
In this work, there are three competing constraints. The tool should (1) return
lenses quickly to users, (2) return the highest quality lenses, and (3)
not require users to write a large number of examples.
Our termination heuristic works well for most situations, but
there are cases where it fails.

@@

In this paper, we have designed and implemented a new, pragmatic
formulation of symmetric lenses as well as new program synthesis
techniques.  In this section, we analyze the relationship
between our simple symmetric
lenses and two previous lens languages:  classical symmetric lenses
and asymmetric lenses.  We also comment on related
program synthesis techniques.

@@

\paragraph*{Symmetric Lenses full Detail}
\begin{centermath}
\begin{array}[b]{l@{\qquad \; \qquad}l}
\begin{array}[b]{c}
\quad\\
  \inferrule*
  {
  }
  {
    \IdentityLensOf{\BRegex} \OfType \BRegex \Leftrightarrow \BRegex
  }
  \\
  \quad
  \end{array}
&
\begin{array}[b]{r@{\ }c@{\ }l}
    \CreateR{} \App s & = & s\\
    \CreateL{} \App s & = & s\\
    \PutR{} \App s \App t & = & s\\
    \PutL{} \App s \App t & = & s\\
  \end{array}
\end{array}
\end{centermath}
Note that the identity lens ignores the second argument in the put functions.
Because the two formats are fully synchronized, no knowledge of the prior data
is needed.

@@

\paragraph*{Symmetric DNF Lenses: Syntax, Cost, Typing, and Semantics}

@@

\section{Forgetful Symmetric Lenses}

\begin{property}[Starting Forgetfulness RL]
  \label{prop:forget-rl}
  Let $\Lens$ be a forgetful symmetric lens.  If $(x_1',c_1') = \Lens.\PutL \App
  (y,\Snd \App (\Lens.\PutR \App (x,c_1)))$, and
  $(x_2',c_2') = \Lens.\PutL \App
  (y,\Snd \App (\Lens.\PutR \App (x,c_2)))$, then $x_1' = x_2'$.
\end{property}
\begin{proof}
  By \ForgetfulRL, $c_1' = c_2'$. We know $(y,c_1') = \Lens.\PutR \App
  (x_1',c_1')$ and $(y,c_1') = \Lens.\PutR \App (x_2',c_1')$ by \PutLR. By
  \PutRL, we know $(x_1' = \Lens.\PutL \App (y,c_1'))$ and $(x_2' = \Lens.\PutL
  \App (y,c_1')))$. Therefore, by transitivity of equality, $x_1' = x_2'$.
\end{proof}

\begin{property}[Starting Forgetfulness LR]
  \label{prop:forget-lr}
  Let $\Lens$ be a forgetful symmetric lens.  If $(y_1',c_1') = \Lens.\PutR \App
  (x,\Snd \App (\Lens.\PutL \App (y,c_1)))$, and
  $(x_2',c_2') = \Lens.\PutL \App
  (x,\Snd \App (\Lens.\PutR \App (y,c_2)))$,
  then $x_1' = x_2'$.
\end{property}
\begin{proof}
  Symmetric to Starting Forgetfulness RL.
\end{proof}

@@

\section{Forgetful Symmetric Lenses}
\begin{theorem*}
  If $\Regex \SSREquiv \RegexAlt$ then $\ProbabilityOf{\Regex}{\String} =
  \ProbabilityOf{\RegexAlt}{\String}$, for all strings $\String \in \LanguageOf{\Regex}$.
\end{theorem*}
The following lemma is
\begin{proof}
\begin{enumerate}

@@

Since $S$ is unambigous, then $S_1$ and $S_2$ are also unambiguous, and for each $s \in \mathcal{L}(S_1 \cdot S_2)$ there exists a unique $s_1 \in \mathcal{L}(S_1)$ and a unique $s_2 \in \mathcal{L}(S_2)$ such that $s = s_1 \cdot s_2$. By the induction hypothesis, $H(\mathcal{L}(S_1)) = \mathbb{H}(S_1)$ and $H(\mathcal{L}(S_2)) = \mathbb{H}(S_2)$, hence\bcp{Don't understand what's going on with the $-$
  signs in the last few lines here.}
\begin{align*}
H(\mathcal{L}(S_1 \cdot S_2))(s) &= -\sum_{s_1 \in S_1} \sum_{s_2 \in S_2}P_{S_1}(s_1)*P_{s_2}(s_2) \log_2(P_{S_1}(s_1)*P_{S_2}(s_2))\\
&= -\sum_{s_1 \in S_1} P_{S_1}(s_1)\sum_{s_2 \in S_2}P_{s_2}(s_2) (\log_2(P_{S_1}(s_1) + \log_2 P_{S_2}(s_2)))\\
&= -\sum_{s_1 \in S_1} P_{S_1}(s_1) \sum_{s_2 \in S_2}P_{S_2}(s_2)\log_2(P_{S_1}(s_1)) + \sum_{s_1 \in S_1} P_{S_1}(s_1)\sum_{s_2 \in S_2} P_{S_2}(s_2)\log_2 P_{S_2}(s_2)\\
&= -\sum_{s_1 \in S_1} P_{S_1}(s_1) \log_2(P_{S_1}(s_1)) + \sum_{s_2 \in S_2} P_{S_2}(s_2)\log_2 P_{S_2}(s_2)\\
&= H(\mathcal{L}(S_1)) + H(\mathcal{L}(S_2))\\
&= \mathbb{H}(S_1) + \mathbb{H}(S_2)
\end{align*}
\item
($S = S_1 \; |_p \; S_2$)

\bcp{``Since...then...'' is not english.}
Since $S$ is unambiguous, then $S_1$ and $S_2$ are also unambiguous, and
every $s \in \mathcal{L}(S_1 \; |_p \; S_2)$ is either in $\mathcal{L}(S_1)$
or in $\mathcal{L}(S_2)$. By the induction hypothesis, $H(\mathcal{L}(S_1))
= \mathbb{H}(S_1)$ and $H(\mathcal{L}(S_2)) =
\mathbb{H}(S_2)$. Then\bcp{Again, don't understand what's going on with the $-$
  signs.}
\begin{align*}
H(\mathcal{L}(S_1 \; |_p \; S_2)) &= -\sum_{s \in \mathcal{L}(S_1)}(p * P_{S_1}(s))\log_2 (p * P_{S_1}(s)) - \sum_{s \in \mathcal{L}(S_2)}((1-p) * P_{S_2}(s))\log_2 ((1-p) * P_{S_2}(s))\\
&=- p\sum_{s \in \mathcal{L}(S_1)}P_{S_1}(s)(\log_2 p + \log_2 P_{S_1}(s)) - (1-p)\sum_{s \in \mathcal{L}(S_2)}P_{S_2}(s)(\log_2 (1-p) + \log_2 P_{S_2}(s))\\
&= p * (\log_2 p +  H(\mathcal{L}(S_1))) + (1-p) * ( \log_2 (1-p) + H(\mathcal{L}(S_2)))\\
&= p * (\log_2 p +  \mathbb{H}(S_1)) + (1-p) * ( \log_2 (1-p) + \mathbb{H}(S_2))\\
\end{align*}
\end{enumerate}
\end{proof}
\begin{theorem*}
  Let $\Lens$ be an asymmetric lens. $\Lens$ is also a simple symmetric lens,
  where
  \begin{center}
    \begin{tabular}{rcl}
      $\Lens.\CreateROf{x}$ & $=$ & $\Lens.get \App x$\\
      $\Lens.\CreateLOf{y}$ & $=$ & $\Lens.create \App y$\\
      $\Lens.\PutROf{x}{y}$ & $=$ & $\Lens.get \App x$\\
      $\Lens.\PutLOf{y}{x}$ & $=$ & $\Lens.put \App y \App x$
    \end{tabular}
  \end{center}
\end{theorem*}
\begin{proof}
Let $\ell$ be an asymmetric lens. Then $\ell$ satisfies the following laws:
\begin{align*}
\ell.\get \; (\ell.\pput \; v \; s) &= v \tag{PUTGET}\\
\ell.\pput \; (\ell.\get \; s) \; s&= s \tag{GETPUT}\\
\ell.\get \; (\ell.\pput \; v \; s) &= v \tag{CREATGET}
\end{align*}
Define the simple symmetric lens $\lceil \ell \rceil$ by 
\begin{align*}
\lceil \ell \rceil.\CreateROf{s} &= \ell.\get \; s\\
\lceil \ell \rceil.\CreateLOf{v} &= \ell.\create \; v\\
\lceil \ell \rceil.\PutROf{s}{v} &= \ell.\get \; s\\
\lceil \ell \rceil.\PutLOf{v}{s} &= \ell.\pput \; v \; s\\
\end{align*}
We now show that $\lceil \ell \rceil$ satisfies the simple symmetric lens laws:
\begin{align*}
\lceil \ell \rceil.\PutLOf{(\lceil \ell \rceil.\CreateROf{s})}{s} &= \ell.\pput \; (\ell.\get \; s) \; s = s\\
\lceil \ell \rceil.\PutROf{(\lceil \ell \rceil.\CreateLOf{v})}{v} &= \ell.\get \; (\ell.\create \; v) = v\\
\lceil \ell \rceil.\PutLOf{(\lceil \ell \rceil.\PutROf{s}{v})}{s} &= \ell.\pput \; (\ell.\get) \; s = s\\
\lceil \ell \rceil.\PutROf{(\lceil \ell \rceil.\PutLOf{v}{s})}{v} &= \ell.\get \; (\ell.\pput \; v \; s) = v
\end{align*}
\end{proof}

>>>>

{\em Lenses} are programs that can be run both ``front to back'' and ``back
to front,'' allowing updates to both input and output data to be transferred
in both directions.
Since their introduction by~\citet{Focal2005-long}, lenses have been extensively
studied, extended, and applied.  Recent work~\cite{optician} has also demonstrated how
techniques from {\em type-directed program synthesis} can be used to
efficiently synthesize a simple class of lenses---so-called {\em
  bijective lenses} over string data---given a pair of types (regular
expressions) and a small number of examples.

\hspace*{1em}We show how to extend this synthesis algorithm to a much wider class of
lenses, which we call {\em simple symmetric lenses}, including not only all
of the bijective lenses and the more widely used ``asymmetric''
lenses~\cite{Focal2005-long}, but also a rich subset of the full-blown
``symmetric lenses'' proposed by \citet{symmetric-lenses}.

@@

In summary, our contributions are:
\begin{itemize}
%
\item We define {\em simple symmetric lenses}, a subclass of symmetric
lenses that do not maintain any hidden state (\S\ref{sec:overview}).  
%
We present a collection of combinators for building simple symmetric lenses and
describe how to calculate their likelihoods (\S\ref{sec:ssl}).
\item We give an algorithm for synthesizing lenses built from these combinators,
  using a novel application of stochastic regular expressions to guide the
  search process (\S\ref{sec:synthesis}) and prove that our metrics have deep
  semantic meaning.
%
\item We extend the Boomerang implementation with simple symmetric lenses,
  demonstrating that they integrate smoothly with other lens extensions such as
  quotient lenses~\cite{quotientlenses} and matching
  lenses~\cite{matchinglenses}. We evaluate our implementation, the effects of
  various optimizations, and our inclusion of relevance annotations on a set of
  48 lens synthesis benchmarks drawn from data cleaning, view maintenance, and
  file synchronization tasks. The system can synthesize simple symmetric lenses
  for all of the benchmarks in under 30 seconds (\S\ref{sec:evaluation}).
%
\item We show that simple symmetric lenses fall between asymmetric
lenses~\cite{Focal2005-long} and full symmetric
lenses~\cite{symmetric-lenses} in expressiveness.  Indeed, the
class of simple lenses can be characterized semantically as the subset of
full symmetric lenses for which a simple ``forgetfulness'' property holds
(\S\ref{sec:related}).
\end{itemize}
Along the
way, we show that {\em star-semiring equivalences} on regular expressions
remain valid when extended to stochastic regular expressions, preserving
probability distributions as well as languages (\S\ref{sec:sre}).
We close with related work (\S\ref{sec:related}) and
concluding thoughts (\S\ref{sec:conc}).
Elided proofs can be found in the appendices
\ifappendices\else included with the supplemental material\fi.

@@

  \setbox0=\hbox{%
    \begin{minipage}{3.5in}
\begin{lstlisting}[
stringstyle=\upshape\sffamily
]
FirstLast,Company
Jane Doe,Healthcare Inc.
John Public,Insurance Co.
\end{lstlisting}
    \end{minipage}
  }
  \savestack{\listingB}{\box0}

@@

  \setbox0=\hbox{%
    \begin{minipage}{3.5in}
\begin{lstlisting}
let company = (name . ("Co." | "Inc." | "Ltd.")) | "UNK" in
let emp_ins = name . " " . name "," company in
let header = "FirstLast,Company" in
let emp_insurance = header . ("\n" . emp_ins)$^*$
\end{lstlisting}
    \end{minipage}
  }
  \savestack{\listingD}{\box0}

  \centering
  \resizebox{\columnwidth}{!}{%
    \begin{tabular}{|>{\columncolor{vlightyellow}}c|>{\columncolor{vlightblue}}c|}
      \hline
      Management's data & HR's data \\
      \hline
      \listingA & \listingB \\
      \hline
      \cellcolor{lightyellow}Management's type & 
      \cellcolor{lightblue} HR's type\\
      \hline
      \cellcolor{lightyellow}\listingC & \cellcolor{lightblue}\listingD \\
      \hline
    \end{tabular}
  }
  \caption{Hypothetical example data files and corresponding regular expressions
  used by management and HR at a US company to represent employee salaries and health
  insurance providers, respectively.
  }
  \label{fig:minimized-representations}
\end{figure}

@@

\begin{theorem}
  If $\Regex \SSREquiv \RegexAlt$ then $\ProbabilityOf{\Regex}{\String} =
  \ProbabilityOf{\RegexAlt}{\String}$, for all strings $\String \in \LanguageOf{\Regex}$.
\end{theorem}
\begin{proof}
  See the appendix %
\ifappendices  Theorem~\ref{thm:semantics-correctness} \else{} included with the supplemental
material\fi.
\end{proof}

@@

We now give typing rules and semantics for (some of) these combinators. We do not
include the semantics for combinators equivalent to those presented in prior
work~\cite{Boomerang}. Furthermore, we present
just \CreateR and \PutR in cases where \CreateL and \PutL are symmetric;
full details can be found in an appendix\ifappendices \S\ref{sec:appendixlenses}  \else{} (in the supplemental materials)\fi. For
simplicity of presentation, we use unambiguous regular expressions
everywhere, even though the identity and disconnect lenses can actually be defined over ambiguous
regular expressions; we believe relaxing
this restriction is unproblematic.

@@

The details of these definitions are syntactically tedious, but not
intellectually difficult. We elide them here but include them in the
appendix \ifappendices \S\ref{sec:appendixdnf}.  \else{} (in the supplemental
materials).\fi

@@

\subsection{Optimizations}
\label{subsec:optimizations}

@@

\subsection{Relevance Annotations}

In this work, there are three competing constraints. The tool should (1) return
lenses quickly to users, (2) return the highest quality lenses, and (3)
not require users to write a large number of examples.
Our termination heuristic works well for most situations, but
there are cases where it fails.

@@

In this paper, we have designed and implemented a new, pragmatic
formulation of symmetric lenses as well as new program synthesis
techniques.  In this section, we analyze the relationship
between our simple symmetric
lenses and two previous lens languages:  classical symmetric lenses
and asymmetric lenses. Proofs about these relationships are included in
the appendix
\ifappendices  \S\ref{sec:appendixforget}. \else{} included with the supplemental
material.\fi

We also comment on related
program synthesis techniques.

@@

\section{Symmetric Lenses full Detail}
\label{sec:appendixlenses}

Here we provide fullly detailed information on symmetric lenses -- full typing
rules and full semantics.

\begin{centermath}
\begin{array}[b]{l@{\qquad \; \qquad}l}
\begin{array}[b]{c}
\quad\\
  \inferrule*
  {
  }
  {
    \IdentityLensOf{\BRegex} \OfType \BRegex \Leftrightarrow \BRegex
  }
  \\
  \quad
  \end{array}
&
\begin{array}[b]{r@{\ }c@{\ }l}
    \CreateR{} \App s & = & s\\
    \CreateL{} \App s & = & s\\
    \PutR{} \App s \App t & = & s\\
    \PutL{} \App s \App t & = & s\\
  \end{array}
\end{array}
\end{centermath}
Note that the identity lens ignores the second argument in the put functions.
Because the two formats are fully synchronized, no knowledge of the prior data
is needed.

@@

\section{Symmetric DNF Lenses: Syntax, Cost, Typing, and Semantics}
\label{sec:appendixdnf}

Here we give full details on symmetric DNF lenses, including syntax, cost,
typing, and semantics.

@@

\section{Forgetful Symmetric Lenses}
\label{sec:appendixforget}

\begin{property}[Starting Forgetfulness RL]
  \label{prop:forget-rl}
  Let $\Lens$ be a forgetful symmetric lens.  If $(x_1',c_1') = \Lens.\PutL \App$
  $(y,\Snd \App (\Lens.\PutR \App (x,c_1)))$, and
  $(x_2',c_2') = \Lens.\PutL \App
  (y,\Snd \App (\Lens.\PutR \App (x,c_2)))$, then $x_1' = x_2'$.
\end{property}
\begin{proof}
  By \ForgetfulRL, $c_1' = c_2'$. We know $(y,c_1') = \Lens.\PutR \App
  (x_1',c_1')$ and $(y,c_1') = \Lens.\PutR \App (x_2',c_1')$ by \PutLR. By
  \PutRL, we know $(x_1' = \Lens.\PutL \App (y,c_1'))$ and $(x_2' = \Lens.\PutL
  \App (y,c_1')))$. Therefore, by transitivity of equality, $x_1' = x_2'$.
\end{proof}

\begin{property}[Starting Forgetfulness LR]
  \label{prop:forget-lr}
  Let $\Lens$ be a forgetful symmetric lens.  If $(y_1',c_1') = \Lens.\PutR \App$
  $(x,\Snd \App (\Lens.\PutL \App (y,c_1)))$, and
  $(x_2',c_2') = \Lens.\PutL \App
  (x,\Snd \App (\Lens.\PutR \App (y,c_2)))$,
  then $x_1' = x_2'$.
\end{property}
\begin{proof}
  Symmetric to Starting Forgetfulness RL.
\end{proof}

@@

\section{SREs and Information Theory}
\begin{theorem}
  \label{thm:semantics-correctness}
  If $\Regex \SSREquiv \RegexAlt$ then $\ProbabilityOf{\Regex}{\String} =
  \ProbabilityOf{\RegexAlt}{\String}$, for all strings $\String \in \LanguageOf{\Regex}$.
\end{theorem}
The following lemma is
\begin{proof}
\begin{enumerate}

@@

Since $S$ is unambigous, then $S_1$ and $S_2$ are also unambiguous, and for each $s \in \mathcal{L}(S_1 \cdot S_2)$ there exists a unique $s_1 \in \mathcal{L}(S_1)$ and a unique $s_2 \in \mathcal{L}(S_2)$ such that $s = s_1 \cdot s_2$. By the induction hypothesis, $H(\mathcal{L}(S_1)) = \mathbb{H}(S_1)$ and $H(\mathcal{L}(S_2)) = \mathbb{H}(S_2)$, hence\bcp{Don't understand what's going on with the $-$
  signs in the last few lines here.}\afm{in the definition of entropy there is a negative. this is to
  address that we're taking the log of a probability, which inherently will
  return a negative number}
\begin{align*}
H(\mathcal{L}(S_1 \cdot S_2))(s) &= -\sum_{s_1 \in S_1} \sum_{s_2 \in S_2}P_{S_1}(s_1)*P_{s_2}(s_2) \log_2(P_{S_1}(s_1)*P_{S_2}(s_2))\\
&= -\sum_{s_1 \in S_1} P_{S_1}(s_1)\sum_{s_2 \in S_2}P_{s_2}(s_2) (\log_2(P_{S_1}(s_1) + \log_2 P_{S_2}(s_2)))\\
&= -\sum_{s_1 \in S_1} P_{S_1}(s_1) \sum_{s_2 \in S_2}P_{S_2}(s_2)\log_2(P_{S_1}(s_1)) + \sum_{s_1 \in S_1} P_{S_1}(s_1)\sum_{s_2 \in S_2} P_{S_2}(s_2)\log_2 P_{S_2}(s_2)\\
&= -\sum_{s_1 \in S_1} P_{S_1}(s_1) \log_2(P_{S_1}(s_1)) + \sum_{s_2 \in S_2} P_{S_2}(s_2)\log_2 P_{S_2}(s_2)\\
&= H(\mathcal{L}(S_1)) + H(\mathcal{L}(S_2))\\
&= \mathbb{H}(S_1) + \mathbb{H}(S_2)
\end{align*}
\item
($S = S_1 \; |_p \; S_2$)

Since $S$ is unambiguous, it must be that $S_1$ and $S_2$ are also unambiguous, and
every $s \in \mathcal{L}(S_1 \; |_p \; S_2)$ is either in $\mathcal{L}(S_1)$
or in $\mathcal{L}(S_2)$. By the induction hypothesis, $H(\mathcal{L}(S_1))
= \mathbb{H}(S_1)$ and $H(\mathcal{L}(S_2)) =
\mathbb{H}(S_2)$. Then\bcp{Again, don't understand what's going on with the $-$
  signs.}
\begin{align*}
H(\mathcal{L}(S_1 \; |_p \; S_2)) &= -\sum_{s \in \mathcal{L}(S_1)}(p * P_{S_1}(s))\log_2 (p * P_{S_1}(s)) - \sum_{s \in \mathcal{L}(S_2)}((1-p) * P_{S_2}(s))\log_2 ((1-p) * P_{S_2}(s))\\
&=- p\sum_{s \in \mathcal{L}(S_1)}P_{S_1}(s)(\log_2 p + \log_2 P_{S_1}(s)) - (1-p)\sum_{s \in \mathcal{L}(S_2)}P_{S_2}(s)(\log_2 (1-p) + \log_2 P_{S_2}(s))\\
&= p * (\log_2 p +  H(\mathcal{L}(S_1))) + (1-p) * ( \log_2 (1-p) + H(\mathcal{L}(S_2)))\\
&= p * (\log_2 p +  \mathbb{H}(S_1)) + (1-p) * ( \log_2 (1-p) + \mathbb{H}(S_2))\\
\end{align*}
\end{enumerate}
\end{proof}
\begin{theorem*}
  Let $\Lens$ be an asymmetric lens. $\Lens$ is also a simple symmetric lens,
  where
  \begin{center}
    \begin{tabular}{rcl}
      $\Lens.\CreateROf{x}$ & $=$ & $\Lens.get \App x$\\
      $\Lens.\CreateLOf{y}$ & $=$ & $\Lens.create \App y$\\
      $\Lens.\PutROf{x}{y}$ & $=$ & $\Lens.get \App x$\\
      $\Lens.\PutLOf{y}{x}$ & $=$ & $\Lens.put \App y \App x$
    \end{tabular}
  \end{center}
\end{theorem*}
\begin{proof}
Let $\ell$ be an asymmetric lens. Then $\ell$ satisfies the following laws:
\begin{align*}
\ell.\get \; (\ell.\pput \; v \; s) &= v \tag{PUTGET}\\
\ell.\pput \; (\ell.\get \; s) \; s&= s \tag{GETPUT}\\
\ell.\get \; (\ell.\pput \; v \; s) &= v \tag{CREATGET}
\end{align*}
Define the simple symmetric lens $\lceil \ell \rceil$ by 
\begin{align*}
\lceil \ell \rceil.\CreateROf{s} &= \ell.\get \; s\\
\lceil \ell \rceil.\CreateLOf{v} &= \ell.\create \; v\\
\lceil \ell \rceil.\PutROf{s}{v} &= \ell.\get \; s\\
\lceil \ell \rceil.\PutLOf{v}{s} &= \ell.\pput \; v \; s\\
\end{align*}
We now show that $\lceil \ell \rceil$ satisfies the simple symmetric lens laws:
\begin{align*}
\lceil \ell \rceil.\PutLOf{(\lceil \ell \rceil.\CreateROf{s})}{s} &= \ell.\pput \; (\ell.\get \; s) \; s = s\\
\lceil \ell \rceil.\PutROf{(\lceil \ell \rceil.\CreateLOf{v})}{v} &= \ell.\get \; (\ell.\create \; v) = v\\
\lceil \ell \rceil.\PutLOf{(\lceil \ell \rceil.\PutROf{s}{v})}{s} &= \ell.\pput \; (\ell.\get) \; s = s\\
\lceil \ell \rceil.\PutROf{(\lceil \ell \rceil.\PutLOf{v}{s})}{v} &= \ell.\get \; (\ell.\pput \; v \; s) = v
\end{align*}
\end{proof}
